GPU-Beschleunigung der Material Point Method

### 2019-01-28 mscthesis.tex
smol fixes

\clearpage
\selectlanguage{english}

\vfill
\begin{center}
\subsubsection*{Zusammenfassung}
\end{center}
Die Material Point Method hat sich in der Computergrafik für physikalische basier\-te Simulationen etabliert. Sie benutzt ein hybrides Modell aus Lagrange-Partikeln und Euler-Gitter. Während Partikel als konsistenter Speicher fungieren, erlaubt das Gitter anfallende partielle Differentialgleichungen effizient zu lösen.

@@

Next up is the transfer: Each thread corresponds to one particle within the blue nodes in Figure \ref{fig:g2p_transfer}. The threads will in unison run over their relative neighbors within the support and collect the weighted contribution to this particle. In this process the particle's position needs to be loaded to the weighted corresponding contribution. The transfer needs to be repeated for each particle in the bin. Since the number of particles is variable, this loop is dynamic and will lead to branch divergence within a warp. One should rely on external mechanics like particle resampling to reduce branching.

@@

Immediately apparent becomes that the SoA-Layout is 2.26 times as fast. However, both memory subunits(VRAM,L2) have the same utilization as in an AoS-Layout. Though the big difference is in the memory access, the wrong access may be conversely spotted in the non-memory subunits. Since an AoS-Layout loads twice as many \texttt{vec4}, every non-memory related metric shrinks by half. An AoS-Layout however issues the same instructions. The SM should do equal work and be active for the same amount of cycles which is not the case.
\begin{table}[htpb]
  \begin{tabular}{ | l | l | l | l | l | l | l |}    \hline
    Layout &  $\Delta t_c (\mu s)$ & Speedup       &VRAM             & L2              & L2-Hit & Read Active \\\hline
    AoS         & 275                   & -        &61.3\%           & \textbf{41.8}\% & \textbf{53.8}\% & 48.9\% \\\hline
    SoA         & 240                   & 1.16x    &75.4\%  & 29.4\%          & 20.0\% & 62.3\% \\
    \hline
  \end{tabular}
\caption{Metrics between different buffer layouts. Map operation with two instructions \texttt{float x = length(vel)}, \texttt{float y = length}\texttt{(pos)} on 1024 $\times$ 1024 \texttt{vec4} buffer storing position(\texttt{pos}) and velocity (\texttt{vel}).}
\label{tab:map_2}
\end{table}

>>>>

\clearpage
\selectlanguage{english}
\vfill
\begin{center}
\subsubsection*{Zusammenfassung}
\end{center}
Die Material Point Method hat sich in der Computergrafik für physikalische basier\-te Simulationen etabliert. Sie benutzt ein hybrides Modell aus Lagrange-Partikeln und Euler-Gitter. Während Partikel als konsistenter Speicher fungieren, erlaubt das Gitter anfallende partielle Differentialgleichungen effizient zu lösen.

@@

Next up is the transfer: Each thread corresponds to one particle within the blue nodes in Figure \ref{fig:g2p_transfer}. The threads run in unison over their relative neighbors within the support and collect the weighted contribution to this particle. In this process the particle's position needs to be loaded to the weighted corresponding contribution. The transfer needs to be repeated for each particle in the bin. Since the number of particles is variable, this loop is dynamic and will lead to branch divergence within a warp. One should rely on external mechanics like particle resampling to reduce branching.

@@

Immediately apparent becomes that the SoA-Layout is 2.26 times faster. However, both memory subunits(VRAM,L2) have the same utilization as in an AoS-Layout. Though the big difference is in the memory access, the wrong access may be conversely spotted in the non-memory subunits. Since an AoS-Layout loads twice as many \texttt{vec4}, every non-memory related metric shrinks by half. An AoS-Layout however issues the same instructions. The SM should do equal work and be active for the same amount of cycles which is not the case.
\begin{table}[htpb]
  \begin{tabular}{ | l | l | l | l | l | l | l |}    \hline
    Layout &  $\Delta t_c (\mu s)$ & Speedup       &VRAM             & L2              & L2-Hit & Read Active \\\hline
    AoS         & 275                   & -        &61.3\%           & \textbf{41.8}\% & \textbf{53.8}\% & 48.9\% \\\hline
    SoA         & 240                   & 1.16x    &75.4\%  & 29.4\%          & 20.0\% & 62.3\% \\
    \hline
  \end{tabular}
\caption{Metrics between different buffer layouts. Map operation with two instructions \texttt{float x = length(vel)}, \texttt{float y = length}\texttt{(pos)} on 1024 $\times$ 1024 \texttt{vec4} buffer storing position(\texttt{pos}) and velocity (\texttt{vel}).}
\label{tab:map_2}
\end{table}
